Data preparation

### 2023-10-18 procesamiento/prod_prep.Rmd
ajuste

```{r eval=FALSE, include=FALSE}
knitr::knit(input = here::here("procesamiento/prod_prep.Rmd"));file.remove("prod_prep.md")
```

@@

frq(df1$oesch16)

df1$oesch16 <- as_labelled(df1$oesch16)
df1$oesch16spo <- as_labelled(df1$oesch16spo)
# # Upper-middle class
#  [1] "'Large employers'"
#  [2] "'Self-employed professionals'"
# 
#  [9] "'Higher-grade managers and administrators'"
# [10] "'Lower-grade managers and administrators'"

@@

df1$oesch3res <- car::recode(df1$oesch16,recode="c(1,2,9,10,5,6,13,14)=1;c(3,4,11)=2;c(7,8,12,15,16)=3")
df1$oesch3spo <- car::recode(df1$oesch16spo,recode="c(1,2,9,10,5,6,13,14)=1;c(3,4,11)=2;c(7,8,12,15,16)=3")
```

@@

```{r}
save(df1,file = here::here("input/data/proc/study1.RData"))
knitr::knit(input = here::here("procesamiento/prod_prep-macro.Rmd"));file.remove("prod_prep-macro.md")
```

>>>>

```{r eval=FALSE, include=FALSE}
# knitr::knit(input = here::here("procesamiento/prod_prep.Rmd"));file.remove("prod_prep.md")
# knitr::knit(input = here::here("procesamiento/prod_prep-macro.Rmd"));file.remove("prod_prep-macro.md")
```

@@

# Traditional Upper-middle class
#  [1] "'Large employers'"
#  [2] "'Self-employed professionals'"
# 
# Salaried middle class
#  [9] "'Higher-grade managers and administrators'"
# [10] "'Lower-grade managers and administrators'"

@@

# reorder Oesch's scheme in a (semi)hierarchical order
oesch_hierar <- 
c("'Large employers'","'Self-employed professionals'",
  "'Higher-grade managers and administrators'",
  "'Lower-grade managers and administrators'",
  "'Technical experts'","'Technicians'",
  "'Socio-cultural professionals'","'Socio-cultural semi-professionals'",
  "'Small business owners with employees'","'Small business owners without employees'","'Skilled clerks'",
  "'Skilled manual'","'Skilled service'",
  "'Low-skilled service'","'Unskilled clerks'","'Low-skilled manual'")

# Respondent social class ------------------------------------------------------
df1$oesch16 <- factor(df1$oesch16,
                      levels = oesch_hierar)
frq(df1$oesch16)

# Partner social class ---------------------------------------------------------
df1$oesch16spo <- factor(df1$oesch16spo,
                      levels = oesch_hierar)
frq(df1$oesch16spo)

#   Given that the Oesch's Scheme consider the vertical (skills) and horizontal (work-logic) as the main distinction between classes
# I will colapse in 4 classes that allow to distinguish between hierarchical differences

# 1. Traditional Upper-middle class
# 2. Salaried middle-class
# 3. Lower middle class
# 4. Working Class

df1$oesch4 <- NA
df1$oesch4[df1$oesch16 %in%c("'Large employers'","'Self-employed professionals'")] <- 4
df1$oesch4[df1$oesch16 %in%c("'Higher-grade managers and administrators'",
                             "'Lower-grade managers and administrators'",
                             "'Technical experts'","'Technicians'",
                             "'Socio-cultural professionals'",
                             "'Socio-cultural semi-professionals'")] <- 3
df1$oesch4[df1$oesch16 %in%c( "'Small business owners with employees'",
                              "'Small business owners without employees'",
                              "'Skilled clerks'")] <- 2

df1$oesch4[df1$oesch16 %in%c("'Low-skilled service'",
                             "'Unskilled clerks'",
                             "'Low-skilled manual'")] <- 1

frq(df1$oesch4) # higher number is the hierarchical order of the four main classes

df1$oesch4spo <- NA
df1$oesch4spo[df1$oesch16spo %in%c("'Large employers'","'Self-employed professionals'")] <- 4
df1$oesch4spo[df1$oesch16spo %in%c("'Higher-grade managers and administrators'",
                             "'Lower-grade managers and administrators'",
                             "'Technical experts'","'Technicians'",
                             "'Socio-cultural professionals'",
                             "'Socio-cultural semi-professionals'")] <- 3
df1$oesch4spo[df1$oesch16spo %in%c( "'Small business owners with employees'",
                              "'Small business owners without employees'",
                              "'Skilled clerks'")] <- 2

df1$oesch4spo[df1$oesch16spo %in%c("'Low-skilled service'",
                             "'Unskilled clerks'",
                             "'Low-skilled manual'")] <- 1
frq(df1$oesch4spo)
#if NA in class of respondent, replace with spouse
df1$oesch4_dom  <- ifelse(is.na(df1$oesch4),df1$oesch4spo,df1$oesch4)
frq(df1$oesch4_dom)
# if NA in respondents it is assumed to be lower than spouse
# now, with the cases in we have information for respondent and spouse
# we replace the value by the highest one

# Replace the highest class by dominance criteria 
# Keep the highest of two values:
df1$oesch4_dom <- pmax(df1$oesch4, #social class respondent
                       df1$oesch4spo,#social class spouse
                       na.rm = T) # if NA, do nothing
frq(df1$oesch4_dom)
df1$oesch4_dom <- as_numeric(df1$oesch4_dom)
frq(df1$oesch4_dom)

# 3-class scheme Oesch
df1$oesch3 <- car::recode(df1$oesch4,"3:4='Upper-middle Class';2='Lower-middle Class';1='Working Class'",
                          as.factor = T,levels = c("Upper-middle Class","Lower-middle Class","Working Class"))
frq(df1$oesch3)
df1$oesch3spo <- car::recode(df1$oesch4spo,"3:4='Upper-middle Class';2='Lower-middle Class';1='Working Class'",
                          as.factor = T,levels = c("Upper-middle Class","Lower-middle Class","Working Class"))
frq(df1$oesch3spo)

df1$oesch3_dom <- car::recode(df1$oesch4_dom,"3:4='Upper-middle Class';2='Lower-middle Class';1='Working Class'",
                          as.factor = T,levels = c("Upper-middle Class","Lower-middle Class","Working Class"))
frq(df1$oesch3_dom)
```

@@

```{r}
save(df1,file = here::here("input/data/proc/study1.RData"))
```
